---
title: "DSCI 101 Project"
author: "Josie Peterburs"
date: "2023-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mdsr)
library(lubridate)

# team id and things
MTeams <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MTeams.csv")
WTeams <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WTeams.csv")

# regular season
MRegularSeasonDetailedResults <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MRegularSeasonDetailedResults.csv")
WRegularSeasonDetailedResults <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WRegularSeasonDetailedResults.csv")

# tourney background
MNCAATourneySeeds <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MNCAATourneySeeds.csv")
WNCAATourneySeeds <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WNCAATourneySeeds.csv")

# regions
MSeasons <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MSeasons.csv")
WSeasons <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WSeasons.csv")

# tournament results
MNCAATourneyDetailedResults <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MNCAATourneyDetailedResults.csv")
WNCAATourneyDetailedResults <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WNCAATourneyDetailedResults.csv")

# other
conferences <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/Conferences.csv")
MTeamConferences <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/MTeamConferences.csv")
WTeamConferences <- read.csv("C:/Users/gigip/Downloads/march-machine-learning-mania-2023/WTeamConferences.csv")
```

## Q1: What team has the most average wins per season for their time as a D1 team since 2003? (Do men and women separately)
Men: Gonzaga with 26.8 wins.
```{r}
# detailed regular season results start in 2003 so find d1 years since 2003
m_time_as_d1 <- MTeams %>% filter(LastD1Season >= 2003) %>% mutate(d1years_since2003 = ifelse(FirstD1Season < 2003, 21, LastD1Season - FirstD1Season + 1))  
  
# find total number of wins for each team per season. detailed results begin in 2003, so there are at most 21 seasons per team.
m_num_wins_per_season <- MRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(numwins = n())

# join tables by TeamID and get average wins (total wins over the years divided by number of years)
m_wins_and_d1_seasons <- m_num_wins_per_season %>% 
  inner_join(m_time_as_d1, by = c("WTeamID" = "TeamID")) %>% select(Season, numwins, TeamName, d1years_since2003) %>% group_by(TeamName) %>% summarize(avg_win_per_season = sum(numwins) / d1years_since2003) %>%  arrange(-avg_win_per_season) %>% unique()

m_wins_and_d1_seasons %>% head(10)
```
Women: Connecticut with 30.1 wins
```{r}
# don't have d1 years for women, so will have to assume all teams have been D1 since 2010 (when the regular season data begins). we want best time averaged over the years, so this hopefully will not affect the major teams
w_time_as_d1 = WTeams %>% mutate(d1years = 2023 - 2010 + 1) 

w_num_wins_per_season <- WRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(numwins = n())

w_wins_and_d1_seasons <- w_num_wins_per_season %>% 
  inner_join(w_time_as_d1, by = c("WTeamID" = "TeamID")) %>% select(Season, numwins, TeamName, d1years) %>% group_by(TeamName) %>% summarize(avg_win_per_season = sum(numwins) / d1years) %>%  arrange(-avg_win_per_season) %>% unique()

# Connecticut is the top team, which makes sense because of their super long win-streak even if some of the teams were not D1 for all 14 seasons.
w_wins_and_d1_seasons %>% head(10)
```


## Q2: My dad always says that free throws win games. Does the team that makes more free throws (by number and by %) in an individual game typically win? In other words, what is the probability that a team wins an individual game given that they made more free throws (count and percentage)?
Men: Count - 0.6768. Percentage - 0.5761.
```{r}
# by number of free throws
m_count_ft <- MRegularSeasonDetailedResults %>% select(Season, WFTM, LFTM) %>% na.omit() %>% mutate(m_team_with_more_ft = ifelse(WFTM > LFTM, 1, 0)) 
m_count_ft %>% summarize(prob = mean(m_team_with_more_ft))

# by percentage of free throws made
m_prct_ft <- MRegularSeasonDetailedResults %>% select(Season, WFTM, WFTA, LFTM, LFTA) %>% na.omit() %>% mutate(m_wteam_prct = WFTM / WFTA, m_lteam_prct = LFTM / LFTA, m_higher_ft_prct = ifelse(m_wteam_prct > m_lteam_prct, 1, 0)) 
m_prct_ft %>% summarize(prob = mean(m_higher_ft_prct, na.rm = TRUE))
```
Women: Count - 0.6623. Percentage - 0.5660.
```{r}
# by number of free throws
w_count_ft <- WRegularSeasonDetailedResults %>% select(Season, WFTM, LFTM) %>% na.omit() %>% mutate(w_team_with_more_ft = ifelse(WFTM > LFTM, 1, 0)) 
w_count_ft %>% summarize(prob = mean(w_team_with_more_ft))

# by percentage of free throws made
w_prct_ft <- WRegularSeasonDetailedResults %>% select(Season, WFTM, WFTA, LFTM, LFTA) %>% na.omit() %>% mutate(w_wteam_prct = WFTM / WFTA, w_lteam_prct = LFTM / LFTA, w_higher_ft_prct = ifelse(w_wteam_prct > w_lteam_prct, 1, 0)) 
w_prct_ft %>% summarize(prob = mean(w_higher_ft_prct, na.rm = TRUE))
```


## Q2.5: Plot FT percentage and winning percentage on a scatter plot. Facet by year and color-code if the team made the March Madness tournament. 
Men:
```{r}
# win percentage
m_win <- MRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(totalwins = n())
m_loss <- MRegularSeasonDetailedResults %>% group_by(Season, LTeamID) %>% summarize(totalloss = n())

m_combined_games <- m_win %>% left_join(m_loss, by = c("WTeamID" = "LTeamID", "Season" = "Season")) %>% mutate(totalgames = totalwins + totalloss, win_prct = totalwins / totalgames) %>% na.omit() %>% select(Season, WTeamID, win_prct)
m_combined_games %>% head()

# freethrows
m_winningteamFT <- MRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(WtotalFTM = sum(WFTM), WtotalFTA = sum(WFTA))

m_losingteamFT <- MRegularSeasonDetailedResults %>% group_by(Season, LTeamID) %>% summarize(LtotalFTM = sum(LFTM), LtotalFTA = sum(LFTA))

m_totalFT <- m_winningteamFT %>% left_join(m_losingteamFT, by = c("WTeamID" = "LTeamID", "Season" = "Season")) %>% na.omit() %>% mutate(totalFTM = WtotalFTM + LtotalFTM, totalFTA = WtotalFTA + LtotalFTA, ft_prct = totalFTM / totalFTA) %>% select(Season, WTeamID, totalFTM, totalFTA, ft_prct) %>% select(Season, WTeamID, ft_prct)
m_totalFT %>% head()

# did the team make the ncaa tournament
m_tourney_teams <- MNCAATourneySeeds %>% filter(Season >= '2003') %>% mutate(made_tourney = "yes")

# combine
m_ft_and_wins <- m_combined_games %>% left_join(m_totalFT, by = c("WTeamID" = "WTeamID", "Season" = "Season")) %>% left_join(m_tourney_teams, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% select(Season, WTeamID, win_prct, ft_prct, made_tourney) 
m_ft_and_wins %>% head()

m_ft_and_wins[is.na(m_ft_and_wins)] <- "no"

m_ft_and_wins %>% ggplot(aes(x = ft_prct, y = win_prct, color = fct_relevel(made_tourney, "yes"))) + geom_point() + xlab("FT %") + ylab("Win %") + labs(color = "Made Tournament?") + facet_wrap(~ Season)

# look at one specific year
m_ft_and_wins %>% filter(Season == '2022') %>% ggplot(aes(x = ft_prct, y = win_prct, color = fct_relevel(made_tourney, "yes"))) + geom_point() + geom_smooth() + xlab("FT %") + ylab("Win %") + labs(color = "Made 2022 Tournament?")
```
Women:
```{r}
# win percentage
w_win <- WRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(totalwins = n())
w_loss <- WRegularSeasonDetailedResults %>% group_by(Season, LTeamID) %>% summarize(totalloss = n())

w_combined_games <- w_win %>% left_join(w_loss, by = c("WTeamID" = "LTeamID", "Season" = "Season")) %>% mutate(totalgames = totalwins + totalloss, win_prct = totalwins / totalgames) %>% na.omit() %>% select(Season, WTeamID, win_prct)
w_combined_games %>% head()

# freethrows
w_winningteamFT <- WRegularSeasonDetailedResults %>% group_by(Season, WTeamID) %>% summarize(WtotalFTM = sum(WFTM), WtotalFTA = sum(WFTA))

w_losingteamFT <- WRegularSeasonDetailedResults %>% group_by(Season, LTeamID) %>% summarize(LtotalFTM = sum(LFTM), LtotalFTA = sum(LFTA))

w_totalFT <- w_winningteamFT %>% left_join(w_losingteamFT, by = c("WTeamID" = "LTeamID", "Season" = "Season")) %>% na.omit() %>% mutate(totalFTM = WtotalFTM + LtotalFTM, totalFTA = WtotalFTA + LtotalFTA, ft_prct = totalFTM / totalFTA) %>% select(Season, WTeamID, totalFTM, totalFTA, ft_prct) %>% select(Season, WTeamID, ft_prct)
w_totalFT %>% head()

# did the team make the ncaa tournament
w_tourney_teams <- WNCAATourneySeeds %>% filter(Season >= '2010') %>% mutate(made_tourney = "yes")

# combine
w_ft_and_wins <- w_combined_games %>% left_join(w_totalFT, by = c("WTeamID" = "WTeamID", "Season" = "Season")) %>% left_join(w_tourney_teams, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% select(Season, WTeamID, win_prct, ft_prct, made_tourney) 
w_ft_and_wins %>% head()

w_ft_and_wins[is.na(w_ft_and_wins)] <- "no"

w_ft_and_wins %>% ggplot(aes(x = ft_prct, y = win_prct, color = fct_relevel(made_tourney, "yes"))) + geom_point() + xlab("FT %") + ylab("Win %") + labs(color = "Made Tournament?") + facet_wrap(~ Season)

# look at one specific year
w_ft_and_wins %>% filter(Season == '2022') %>% ggplot(aes(x = ft_prct, y = win_prct, color = fct_relevel(made_tourney, "yes"))) + geom_point() + geom_smooth() + xlab("FT %") + ylab("Win %") + labs(color = "Made 2022 Tournament?")
```


## Q3: Was there a difference in score, fouls, assists, and three point shooting percentage by conference in the 21-22 season? Plot each element on sied-by-side boxplots to compare conferences.
(Top conferences from WarrenNolan.com; 2022 season for men and women)

 - Men: West Coast, Atlantic 10, Mountain West, American Athletic, Pac-12, ACC, Big East, SEC, Big Ten, Big 12
 
 - Women: Missouri Valley, Atlantic 10, Big East, West Coast, American Athletic, Big Ten, ACC, Big 12, Pac-12, SEC 

Men (Top 10 Conferences):
```{r}
m_topconf <- c("West Coast Conference", "Atlantic 10 Conference", "Mountain West Conference", "American Athletic Conference", "Pacific-12 Conference", "Atlantic Coast Conference", "Big East Conference", "Southeastern Conference", "Big Ten Conference", "Big 12 Conference")

m_conferences10 <- conferences %>% filter(Description %in% m_topconf)
m_confabbrevs <- m_conferences10 %>% select(ConfAbbrev)
m_confabbrevs2 <- c("a_ten", "aac", "acc", "big_east", "big_ten", "big_twelve", "mwc", "pac_twelve", "sec", "wcc")

m_teams <- MTeamConferences %>% filter(ConfAbbrev %in% m_confabbrevs2, Season == "2022") %>% select(TeamID, ConfAbbrev)

m_win_gamedetails <- MRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(WTeamID, WScore, WPF, WAst, WFGM3, WFGA3) %>% inner_join(m_teams, by = c("WTeamID" = "TeamID")) %>% left_join(m_conferences10, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "WTeamID",
               "Score" = "WScore",
               "Fouls" = "WPF",
               "Assists" = "WAst",
               "FGM3" = "WFGM3",
               "FGA3" = "WFGA3")

m_lose_gamedetails <- MRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(LTeamID, LScore, LPF, LAst, LFGM3, LFGA3) %>% inner_join(m_teams, by = c("LTeamID" = "TeamID")) %>% left_join(m_conferences10, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "LTeamID",
               "Score" = "LScore",
               "Fouls" = "LPF",
               "Assists" = "LAst",
               "FGM3" = "LFGM3",
               "FGA3" = "LFGA3")

m_all_gamedetails <- rbind(m_win_gamedetails, m_lose_gamedetails) %>% mutate(Three_Point_Prct = FGM3 / FGA3)

# Score
m_all_gamedetails %>% ggplot(aes(x = Description, y = Score)) + geom_boxplot() + coord_flip()

# Fouls
m_all_gamedetails %>% ggplot(aes(x = Description, y = Fouls)) + geom_boxplot() + coord_flip()

# Assists
m_all_gamedetails %>% ggplot(aes(x = Description, y = Assists)) + geom_boxplot() + coord_flip()

# 3-Point Shooting % 
m_all_gamedetails %>% ggplot(aes(x = Description, y = Three_Point_Prct)) + geom_boxplot() + coord_flip()

# overall averages
m_category_avgs <- m_all_gamedetails %>% group_by(Description) %>% summarize(score = mean(Score), fouls = mean(Fouls), assists = mean(Assists), threept = mean(FGM3 / FGA3))
m_category_avgs
```
Women (Top Ten Conferences):
```{r}
w_topconf <- c("Missouri Valley Conference", "Atlantic 10 Conference", "Big East Conference", "West Coast Conference", "American Athletic Conference", "Big Ten Conference", "Atlantic Coast Conference", "Big 12 Conference", "Pacific-12 Conference", "Southeastern Conference")

w_conferences10 <- conferences %>% filter(Description %in% w_topconf)
w_confabbrevs <- w_conferences10 %>% select(ConfAbbrev)
w_confabbrevs2 <- c("a_ten", "aac", "acc", "big_east", "big_ten", "big_twelve", "mvc", "pac_twelve", "sec", "wcc")

w_teams <- WTeamConferences %>% filter(ConfAbbrev %in% w_confabbrevs2, Season == "2022") %>% select(TeamID, ConfAbbrev)

w_win_gamedetails <- WRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(WTeamID, WScore, WPF, WAst, WFGM3, WFGA3) %>% inner_join(w_teams, by = c("WTeamID" = "TeamID")) %>% left_join(w_conferences10, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "WTeamID",
               "Score" = "WScore",
               "Fouls" = "WPF",
               "Assists" = "WAst",
               "FGM3" = "WFGM3",
               "FGA3" = "WFGA3")

w_lose_gamedetails <- WRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(LTeamID, LScore, LPF, LAst, LFGM3, LFGA3) %>% inner_join(w_teams, by = c("LTeamID" = "TeamID")) %>% left_join(w_conferences10, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "LTeamID",
               "Score" = "LScore",
               "Fouls" = "LPF",
               "Assists" = "LAst",
               "FGM3" = "LFGM3",
               "FGA3" = "LFGA3")

w_all_gamedetails <- rbind(w_win_gamedetails, w_lose_gamedetails) %>% mutate(Three_Point_Prct = FGM3 / FGA3)

# Score
w_all_gamedetails %>% ggplot(aes(x = Description, y = Score)) + geom_boxplot() + coord_flip()

# Fouls
w_all_gamedetails %>% ggplot(aes(x = Description, y = Fouls)) + geom_boxplot() + coord_flip()

# Assists
w_all_gamedetails %>% ggplot(aes(x = Description, y = Assists)) + geom_boxplot() + coord_flip()

# 3-Point Shooting % 
w_all_gamedetails %>% ggplot(aes(x = Description, y = Three_Point_Prct)) + geom_boxplot() + coord_flip()

# overall averages
w_category_avgs <- w_all_gamedetails %>% group_by(Description) %>% summarize(score = mean(Score), fouls = mean(Fouls), assists = mean(Assists), threept = mean(FGM3 / FGA3))
w_category_avgs
```

Men (all conferences):
```{r}
m_teamsALL <- MTeamConferences %>% filter(Season == "2022") %>% select(TeamID, ConfAbbrev)

m_win_gamedetailsALL <- MRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(WTeamID, WScore, WPF, WAst, WFGM3, WFGA3) %>% inner_join(m_teamsALL, by = c("WTeamID" = "TeamID")) %>% left_join(conferences, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "WTeamID",
               "Score" = "WScore",
               "Fouls" = "WPF",
               "Assists" = "WAst",
               "FGM3" = "WFGM3",
               "FGA3" = "WFGA3")

m_lose_gamedetailsALL <- MRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(LTeamID, LScore, LPF, LAst, LFGM3, LFGA3) %>% inner_join(m_teamsALL, by = c("LTeamID" = "TeamID")) %>% left_join(conferences, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "LTeamID",
               "Score" = "LScore",
               "Fouls" = "LPF",
               "Assists" = "LAst",
               "FGM3" = "LFGM3",
               "FGA3" = "LFGA3")

m_all_gamedetailsALL <- rbind(m_win_gamedetailsALL, m_lose_gamedetailsALL) %>% mutate(Three_Point_Prct = FGM3 / FGA3)

# Score
m_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Score)) + geom_boxplot() + coord_flip()

# Fouls
m_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Fouls)) + geom_boxplot() + coord_flip()

# Assists
m_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Assists)) + geom_boxplot() + coord_flip()

# 3-Point Shooting % 
m_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Three_Point_Prct)) + geom_boxplot() + coord_flip()

# overall averages
m_category_avgsALL <- m_all_gamedetailsALL %>% group_by(Description) %>% summarize(score = mean(Score), fouls = mean(Fouls), assists = mean(Assists), threept = mean(FGM3 / FGA3))
m_category_avgsALL
```
Women (all conferences):
```{r}
w_teamsALL <- WTeamConferences %>% filter(Season == "2022") %>% select(TeamID, ConfAbbrev)

w_win_gamedetailsALL <- WRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(WTeamID, WScore, WPF, WAst, WFGM3, WFGA3) %>% inner_join(w_teamsALL, by = c("WTeamID" = "TeamID")) %>% left_join(conferences, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "WTeamID",
               "Score" = "WScore",
               "Fouls" = "WPF",
               "Assists" = "WAst",
               "FGM3" = "WFGM3",
               "FGA3" = "WFGA3")

w_lose_gamedetailsALL <- WRegularSeasonDetailedResults %>% filter(Season == "2022") %>% select(LTeamID, LScore, LPF, LAst, LFGM3, LFGA3) %>% inner_join(w_teamsALL, by = c("LTeamID" = "TeamID")) %>% left_join(conferences, by = c("ConfAbbrev" = "ConfAbbrev")) %>% 
        rename("TeamID" = "LTeamID",
               "Score" = "LScore",
               "Fouls" = "LPF",
               "Assists" = "LAst",
               "FGM3" = "LFGM3",
               "FGA3" = "LFGA3")

w_all_gamedetailsALL <- rbind(w_win_gamedetailsALL, w_lose_gamedetailsALL) %>% mutate(Three_Point_Prct = FGM3 / FGA3)

# Score
w_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Score)) + geom_boxplot() + coord_flip()

# Fouls
w_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Fouls)) + geom_boxplot() + coord_flip()

# Assists
w_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Assists)) + geom_boxplot() + coord_flip()

# 3-Point Shooting % 
w_all_gamedetailsALL %>% ggplot(aes(x = Description, y = Three_Point_Prct)) + geom_boxplot() + coord_flip()

# overall averages
w_category_avgsALL <- w_all_gamedetailsALL %>% group_by(Description) %>% summarize(score = mean(Score), fouls = mean(Fouls), assists = mean(Assists), threept = mean(FGM3 / FGA3))
w_category_avgsALL
```


## Q4: Which team is more likely to win an overtime game during the regular season (home or away)?
Men: Home team won overtime games 56.17% of the time.
```{r}
m_homewins <- MRegularSeasonDetailedResults %>% filter(NumOT > 0, WLoc == "H") %>% summarize(homecount = n())
m_homewins
m_awaywins <- MRegularSeasonDetailedResults %>% filter(NumOT > 0, WLoc == "A") %>% summarize(awaycount = n())
m_awaywins

m_homewins / (m_homewins + m_awaywins)
```
Women: Home team won overtime games 53.98% of the time.
```{r}
w_homewins <- WRegularSeasonDetailedResults %>% filter(NumOT > 0, WLoc == "H") %>% summarize(homecount = n())
w_homewins
w_awaywins <- WRegularSeasonDetailedResults %>% filter(NumOT > 0, WLoc == "A") %>% summarize(awaycount = n())
w_awaywins

w_homewins / (w_homewins + w_awaywins)
```


## Q5: What is the lowest seed to have won the championship game? What about getting to the Final Four?
Men: The lowest seed to have won the championship in the provided years is Connecticut in 2014 as a 7 seed. The lowest seeds to have made it to the final four are George Mason in 2006, VCU in 2011, Loyola-Chicago in 2018, and UCLA in 2021, all 11 seeds.
```{r}
# won champsionship
m_championshipgames <- MNCAATourneyDetailedResults %>% arrange(-DayNum) %>% filter(DayNum == 154) %>% select(Season, WTeamID) %>% left_join(MNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(MTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName)
m_championshipgames

# made it to final four
m_finalfourW <- MNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, WTeamID, DayNum) %>% filter(DayNum == 152) %>% left_join(MNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(MTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName) %>% rename("TeamID" = "WTeamID")
m_finalfourL <- MNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, LTeamID, DayNum) %>% filter(DayNum == 152) %>% left_join(MNCAATourneySeeds, by = c("Season" = "Season", "LTeamID" = "TeamID")) %>% left_join(MTeams, by = c("LTeamID" = "TeamID")) %>% select(Season, LTeamID, Seed, TeamName) %>% rename("TeamID" = "LTeamID")

m_allfinal4 <- rbind(m_finalfourW, m_finalfourL) %>% arrange(Season)
m_allfinal4
```
Women: The lowest seed to have won the championship in the provided years is Texas A&M in 2011 as a 2 seed. All other championships have been won by a 1 seed. The lowest seed to have made it to the final four is Washington in 2016 as a 7 seed.
```{r}
# championship game was day155 for 2010-2016, but day153 for 2017-2022
w_championshipgames1 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, WTeamID, DayNum) %>% filter(DayNum == 155) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(WTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName) 
w_championshipgames2 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, WTeamID, DayNum) %>% filter(Season %in% c(2017:2022), DayNum == 153) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(WTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName) 
w_championshipgames <- rbind(w_championshipgames1, w_championshipgames2) %>% arrange(Season)
w_championshipgames

# made it to final four
# final four was day 153 for 2010-2016
w_finalfourW1 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, WTeamID, DayNum) %>% filter(Season <= 2016, DayNum == 153) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(WTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName) %>% rename("TeamID" = "WTeamID")

w_finalfourL1 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, LTeamID, DayNum) %>% filter(Season <= 2016, DayNum == 153) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "LTeamID" = "TeamID")) %>% left_join(WTeams, by = c("LTeamID" = "TeamID")) %>% select(Season, LTeamID, Seed, TeamName) %>% rename("TeamID" = "LTeamID")
# 2010-2016
w_allfinal41 <- rbind(w_finalfourW1, w_finalfourL1) %>% arrange(Season)

# final four was day151 for 2017-2022
w_finalfourW2 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, WTeamID, DayNum) %>% filter(Season > 2016, DayNum == 151) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "WTeamID" = "TeamID")) %>% left_join(WTeams, by = c("WTeamID" = "TeamID")) %>% select(Season, WTeamID, Seed, TeamName) %>% rename("TeamID" = "WTeamID")

w_finalfourL2 <- WNCAATourneyDetailedResults %>% arrange(-DayNum) %>% select(Season, LTeamID, DayNum) %>% filter(Season > 2016, DayNum == 151) %>% left_join(WNCAATourneySeeds, by = c("Season" = "Season", "LTeamID" = "TeamID")) %>% left_join(WTeams, by = c("LTeamID" = "TeamID")) %>% select(Season, LTeamID, Seed, TeamName) %>% rename("TeamID" = "LTeamID")
# 2017-2022
w_allfinal42 <-  rbind(w_finalfourW2, w_finalfourL2) %>% arrange(Season)

# all final four teams
w_allfinal4 <- rbind(w_allfinal41, w_allfinal42) %>% arrange(Season)
w_allfinal4
```


## Chip's Question: Compare freethrow percentages between years. Has the average freethrow percentage increased, decreased, or stayed about the same?

Men:
```{r}
# added winning team ft made to losing team ft made. divided by winning and losing attempted ft sum to get overall ft percentage for each game. 
game_ft_perct <- MRegularSeasonDetailedResults %>% select(Season, WFTM, WFTA, LFTM, LFTA) %>% mutate(ft_perct = (WFTM + LFTM) / (WFTA + LFTA))
game_ft_perct %>% head()

season_avg <- game_ft_perct %>% group_by(Season) %>% summarize(avg_ft_perct_year = mean(ft_perct))
season_avg
season_avg %>% ggplot(aes(x = Season, y = avg_ft_perct_year)) + geom_line()

# so ft percentage did increase, but only by 2%
```

Women:
```{r}
game_ft_perct <- WRegularSeasonDetailedResults %>% select(Season, WFTM, WFTA, LFTM, LFTA) %>% mutate(ft_perct = (WFTM + LFTM) / (WFTA + LFTA))
game_ft_perct %>% head()

season_avg <- game_ft_perct %>% group_by(Season) %>% summarize(avg_ft_perct_year = mean(ft_perct))
season_avg
season_avg %>% ggplot(aes(x = Season, y = avg_ft_perct_year)) + geom_line()
```

